---
title: "p8105_hw6_ms6360"
author: "Maisie Sun"
date: "2022-11-21"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
library(patchwork)
library(readxl)
library(viridis)
library(rstatix)
library(PerformanceAnalytics)
library(modelr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r}
weather_df %>% 
  modelr::bootstrap(n = 1000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::glance)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  ggplot(aes(x = r.squared)) + geom_density()
```

```{r}
weather_df %>% 
  modelr::bootstrap(n = 1000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::tidy)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  select(id = `.id`, term, estimate) %>% 
  pivot_wider(
    names_from = term, 
    values_from = estimate) %>% 
  rename(beta0 = `(Intercept)`, beta1 = tmin) %>% 
  mutate(log_b0b1 = log(beta0 * beta1)) %>% 
  ggplot(aes(x = log_b0b1)) + geom_density()
```

# Problem 2

1. Import data

```{r import_data}
homicide_data = 
  read_csv("./data/homicide-data.csv") %>%
  janitor::clean_names()
```

2. Clean data

```{r clean_data}
homicide_data_p2 = 
  homicide_data %>%
  unite("city_state", city:state, remove = FALSE, sep = ", ") %>%
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")) %>%
  filter(victim_race == "White" | victim_race == "Black") %>%
  mutate(resolved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age),
    victim_race = fct_relevel(victim_race, "White")) %>%
  select(city_state, resolved, victim_age, victim_sex, victim_race) %>%
  drop_na()
```

* There are `r nrow(homicide_data_p2)` observations and `r ncol(homicide_data_p2)` variables in this cleaned dataset. The variables are: `r names(homicide_data_p2)`.

```{r logistic_Baltimore}
fit_logistic_Baltimore = 
  homicide_data_p2 %>% 
  filter(city_state == "Baltimore, MD") %>%
  glm(resolved ~ victim_age + victim_race + victim_sex, data = ., family = binomial()) %>%
  broom::tidy(conf.int = TRUE)
```

3. Fit logistic regression for Baltimore, MD, for resolved vs unresolved cases comparing male victims to female victims

```{r baltimore_result}
result_Baltimore = 
  fit_logistic_Baltimore %>%
  mutate(
    OR = exp(estimate),
    lower95CI = exp(conf.low),
    upper95CI = exp(conf.high)) %>%
  select(term, OR, lower95CI, upper95CI) %>%
  filter(term == "victim_sexMale")

result_Baltimore %>%
  knitr::kable(digits = 3)
```

* The estimated odds ratio of resolved homicides comparing male victims to female victims is `r result_Baltimore[,"OR"]`, the 95% confidence interval is (`r result_Baltimore[,"lower95CI"]`, `r result_Baltimore[,"upper95CI"]`). 

4. Fit logistic regression for all cities for resolved vs unresolved cases comparing male victims to female victims

```{r}
results_cities = 
  homicide_data_p2 %>%
  nest(data = -city_state) %>% 
  mutate(
    logistic = map(data, ~glm(resolved ~ victim_age + victim_race + victim_sex, data = ., family = binomial())),
    results = map(.x = logistic, ~broom::tidy(.x,conf.int = TRUE))) %>% 
  select(city_state, results) %>% 
  unnest(results) %>% 
  mutate(
OR = exp(estimate),
    lower95CI = exp(conf.low),
    upper95CI = exp(conf.high)) %>%
  select(city_state, term, OR, lower95CI, upper95CI) %>%
  filter(term == "victim_sexMale")

results_cities %>% 
  knitr::kable(digits = 3)

```

5. Plot showing the estimated ORs and CIs for each city. 

```{r plot}
results_cities %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = lower95CI, ymax = upper95CI), width = .8) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(
    title = "Odds ratio and 95%CI of resolved crime in each city in the US"
  )

```

Resolved homicides OR comparing male and female victims is lowest in New York and highest in Albuquerque. Albuquerque also has the largest 95%CI, which means there is large amount of variation within the data for Albuquerque.

## Problem 3

1. Import data

```{r data_import}
birthweight_data = 
  read_csv("./data/birthweight.csv") %>%
  janitor::clean_names()
```

2. Clean data

```{r}
birthweight_data_clean = 
  birthweight_data %>%
  mutate(across(.cols = c(babysex, frace, malform, mrace), as.factor)) %>%
  mutate(
      babysex = as.factor(babysex),
      malform = as.factor(malform),
      frace = as.factor(frace),
      mrace = as.factor(mrace))
```

3. Checking for missing data

```{r, warning = FALSE, message=FALSE}
birthweight_data_clean[!complete.cases(birthweight_data_clean),]  
```

* There is no missing data in this dataset. There are `r nrow(birthweight_data_clean)` observations and `r ncol(birthweight_data_clean)` variables in this cleaned dataset. The variables are: `r names(birthweight_data_clean)`.

4. Output a correlation matrix to see which variables are associated with birthweight

```{r correlation_matrix}
birthweight_data %>% 
  cor_mat() %>%
  cor_gather() %>%
  filter(var1 %in% "bwt", !var2 %in% "bwt") %>%
  mutate(
    cor_if_sig = ifelse(p < 0.01, cor, NA)
    ) %>% 
  ggplot(aes(x = var1, y = var2, fill = cor, label = round(cor_if_sig, 2))) + 
  geom_tile(color = "white") +   
  geom_text(color = "white", size = 4) + 
  labs(
    x = "Birthweight", 
    y = "Predictors",
    title = "Correlations between predictors and birthweight",
    subtitle = "significant predictors with p-value < 0.01",
    fill = "Correlation"
  )
```

* According to the correlation matrix, variables `wtgain`, `ppwt`, `mrace`, `momage`, `mheight`, `gaweeks`, `frace`, `fincome`, `delwt`, `blength`, and `bhead` should be included in the model. These variables have correlation that is larger than 0.1 whether positive or negative.

5. Fit a linear model based on variables associated with birthweight

```{r linear_model}
linear_fit = 
  birthweight_data %>%
  select(bwt, wtgain, ppwt, mrace, momage, mheight, gaweeks, frace, fincome, delwt, blength, bhead)

model_1 = 
  lm(bwt ~ blength + bhead + wtgain + ppwt + mrace + momage + mheight + gaweeks + frace + fincome + delwt, data = linear_fit)

```

6. Plotting fitted values and residuals

```{r}
  linear_fit %>% 
  modelr::add_residuals(model_1) %>%
  modelr::add_predictions(model_1) %>%
  ggplot(aes(x = pred, y = resid)) +
  geom_point() +
  labs(
    x = "Fitted values",
    y = "Residuals",
    title = "Model residuals vs fitted values"
  )
```

7. Comparing models

```{r}
model_2 = lm(bwt ~ blength + gaweeks, data = birthweight_data)
model_3 = lm(bwt ~ bhead * blength * babysex, data = birthweight_data)
```

8. Cross validation

```{r}
cross_validation = 
  crossv_mc(birthweight_data, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

rmse = 
  cross_validation %>% 
  mutate(
   model_1 = map(.x = train, ~lm(bwt ~ blength + bhead + wtgain + ppwt + mrace + momage + mheight + gaweeks + frace + fincome + delwt, data = .x)),
   model_2 = map(.x = train, ~lm(bwt ~ blength + gaweeks, data = .x)),
   model_3 = map(.x = train, ~lm(bwt ~ bhead * blength * babysex, data = .x))
  ) %>% 
  mutate(
    rmse_model1 = map2_dbl(.x = model_1, .y = test, ~rmse(model = .x, data = .y)),
    rmse_model2 = map2_dbl(.x = model_2, .y = test, ~rmse(model = .x, data = .y)),
    rmse_model3 = map2_dbl(.x = model_3, .y = test, ~rmse(model = .x, data = .y))
  )
```

9. Violon plots comparing RMSE for the three models.

```{r}
rmse %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

From the RMSE plot, it appears that of the model 1 has the lowest RMSE out of the 3 models, which means that it is the best fitted model. With more confounders included, the likelihood of getting a significant result is also more likely. However, RMSE from all models are high, therefore, there should be better models to represent the data other than these 3 models.

Model 1 contains variables `wtgain`, `ppwt`, `mrace`, `momage`, `mheight`, `gaweeks`, `frace`, `fincome`, `delwt`, `blength`, and `bhead`. 
