---
title: "p8105_hw6_ms6360"
author: "Maisie Sun"
date: "2022-11-21"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
library(patchwork)
library(readxl)
library(viridis)
library(rstatix)
library(PerformanceAnalytics)
library(modelr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 2

```{r import_data}
homicide_data = 
  read_csv("./data/homicide-data.csv") %>%
  janitor::clean_names()
```

```{r clean_data}
homicide_data_p2 = 
  homicide_data %>%
  unite("city_state", city:state, remove = FALSE, sep = ", ") %>%
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")) %>%
  filter(victim_race == "White" | victim_race == "Black") %>%
  mutate(resolved = case_when(
    startsWith(disposition, "Closed") ~ 1,
    startsWith(disposition, "Open") ~ 0
    ),
    victim_age = as.numeric(victim_age),
    victim_race = fct_relevel(victim_race, "White"))

```

```{r logistic_Baltimore}
fit_logistic_Baltimore = 
  homicide_data_p2 %>% 
  filter(city_state == "Baltimore, MD") %>%
  glm(resolved ~ victim_age + victim_race + victim_sex, data = ., family = binomial()) %>%
  broom::tidy(conf.int = TRUE)

```

```{r baltimore_result}
result_Baltimore = 
  fit_logistic_Baltimore %>%
  mutate(
    OR = exp(estimate),
    lower95CI = exp(conf.low),
    upper95CI = exp(conf.high)) %>%
  select(term, OR, lower95CI, upper95CI) %>%
  filter(term == "victim_sexMale") %>%
  knitr::kable(digits = 3)
```

```{r}
results_cities = 
  homicide_data_p2 %>%
  nest(data = -city_state) %>% 
  mutate(
    logistic = map(data, ~glm(resolved ~ victim_age + victim_race + victim_sex, data = ., family = binomial())),
    results = map(.x = logistic, ~broom::tidy(.x,conf.int = TRUE))) %>% 
  select(city_state, results) %>% 
  unnest(results) %>% 
  mutate(
OR = exp(estimate),
    lower95CI = exp(conf.low),
    upper95CI = exp(conf.high)) %>%
  select(city_state, term, OR, lower95CI, upper95CI) %>%
  filter(term == "victim_sexMale")

results_cities %>% 
  knitr::kable(digits = 3)

```

```{r plot}
results_cities %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = lower95CI, ymax = upper95CI), width = .8) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(
    title = "Odds ratio and 95%CI of resolved crime in each city in the US"
  )

```

## Problem 3

```{r data_import}
birthweight_data = 
  read_csv("./data/birthweight.csv") %>%
  janitor::clean_names()
```

### Clean data

```{r}
birthweight_data_clean = 
  birthweight_data %>%
  mutate(across(.cols = c(babysex, frace, malform, mrace), as.factor)) %>%
  mutate(
      babysex = ifelse(babysex == "1", "male","female"),
      malform = ifelse(malform == "0", "absent","present"),
      frace = recode(frace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other", "9" = "Unknown"),
      mrace = recode(mrace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other")
         )
```

### Checking for missing data
```{r, warning = FALSE, message=FALSE}
birthweight_data_clean[!complete.cases(birthweight_data_clean),]  
```

* There is no missing data in this dataset. There are `r nrow(birthweight_data_clean)` observations and `r ncol(birthweight_data_clean)` variables in this cleaned dataset. The variables are: `r names(birthweight_data_clean)`.

```{r correlation_matrix}
birthweight_data %>% 
  cor_mat() %>%
  cor_gather() %>%
  filter(var1 %in% "bwt", !var2 %in% "bwt") %>%
  mutate(
    cor_if_sig = ifelse(p < 0.01, cor, NA)
    ) %>% 
  ggplot(aes(x = var1, y = var2, fill = cor, label = round(cor_if_sig, 2))) + 
  geom_tile(color = "white") +   
  geom_text(color = "white", size = 4) + 
  labs(
    x = "Birthweight", 
    y = "Predictors",
    title = "Correlations between predictors and birthweight",
    subtitle = "significant predictors with p-value < 0.01",
    fill = "Correlation"
  )
```

* According to the correlation matrix, variables `wtgain`, `ppwt`, `mrace`, `momage`, `mheight`, `gaweeks`, `frace`, `fincome`, `delwt`, `blength`, and `bhead` should be included in the model. These variables have correlation that is larger than 0.1 whether positive or negative.

```{r linear_model}
linear_fit = 
  birthweight_data %>%
  select(bwt, wtgain, ppwt, mrace, momage, mheight, gaweeks, frace, fincome, delwt, blength, bhead)

model_1 = 
  lm(bwt ~ blength + bhead + wtgain + ppwt + mrace + momage + mheight + gaweeks + frace + fincome + delwt, data = linear_fit)

```

```{r}
  linear_fit %>% 
  modelr::add_residuals(model_1) %>%
  modelr::add_predictions(model_1) %>%
  ggplot(aes(x = pred, y = resid)) +
  geom_point() +
  labs(
    x = "Fitted values",
    y = "Residuals",
    title = "Model residuals vs fitted values"
  )
```

### Comparing models

One using length at birth and gestational age as predictors (main effects only)
One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

```{r}
model_2 = lm(bwt ~ blength + gaweeks, data = birthweight_data)
model_3 = lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex + bhead * blength * babysex, data = birthweight_data)
```

#### Cross validation

```{r}
cross_validation = 
  crossv_mc(birthweight_data, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

rmse = 
  cross_validation %>% 
  mutate(
   model_1 = map(.x = train, ~lm(bwt ~ blength + bhead + wtgain + ppwt + mrace + momage + mheight + gaweeks + frace + fincome + delwt, data = .x)),
   model_2 = map(.x = train, ~lm(bwt ~ blength + gaweeks, data = .x)),
   model_3 = map(.x = train, ~lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex + bhead * blength * babysex, data = .x))
  ) %>% 
  mutate(
    rmse_model1 = map2_dbl(.x = model_1, .y = test, ~rmse(model = .x, data = .y)),
    rmse_model2 = map2_dbl(.x = model_2, .y = test, ~rmse(model = .x, data = .y)),
    rmse_model3 = map2_dbl(.x = model_3, .y = test, ~rmse(model = .x, data = .y))
  )
```

Violon plots

```{r}
rmse %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

From the RMSE plot, it appears that of the model 1 has the lowest RMSE, which means that it is the best fitted model. With more confounders included, the likelihood of getting a significant result is also more likely.

Model 1 contains variables `wtgain`, `ppwt`, `mrace`, `momage`, `mheight`, `gaweeks`, `frace`, `fincome`, `delwt`, `blength`, and `bhead`. 
